---
title: "hw02_dplyr"
output: html_document
---
 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Packages

Load tidyverse and gapminder dataset:

```{r load}
library(tidyverse)
library(gapminder)
```

## 1.1

Filtered gapminder for data in only the 1970's:
```{r}
gapminder %>%
  filter(year == 1970:1979)
```

Picked 3 countries within the filtered results and refiltered for Greece, Austria, and Philippines within the 1970's.  Assigned new table to varaible gapminder1.1:
```{r }
gapminder1.1 <- (gapminder %>%
  filter(country == 'Greece' | country == 'Austria' | country == 'Philippines', year == 1970:1979))
gapminder1.1
```

## 1.2

Show variables country and gdpPercap only from gapminder1.1:
```{r}
gapminder1.1 %>%
  select(country, gdpPercap)
```

## 1.3**

specific to each country?  or mush them all together (below)?
```{r}
gapminder %>%
  arrange(gapminder, year) %>%
  filter(diff(gapminder$lifeExp) < 0)
```
```{r}
a <- c(1, 2, 4, 3, 9, 0)
diff(a)
```

## 1.4

To filter gapminder to find the rows with the 3 largest and 3 smallest GDP per capita you can:

Find the number of rows in gapminder:
```{r}
nrow(gapminder)
```
 
Arrange gapminder in descending order for GDP per capita, then select top 3 rows and bottom 3 rows:

```{r}
(gapminder %>%
  arrange(desc(gdpPercap)))[c(1:3, 1702:1704), ]
```

