---
title: "hw02_dplyr"
output: html_document
---
 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Packages

Load tidyverse, gapminder, and datasets:

```{r load}
library(tidyverse)
library(gapminder)
library(datasets)
```

## 1.1

Filtered gapminder for data in only the 1970's:
```{r}
gapminder %>%
  filter(year == 1970:1979)
```

Picked 3 countries within the filtered results and refiltered for Greece, Austria, and Philippines within the 1970's.  Assigned new table to varaible gapminder1.1:
```{r }
gapminder1.1 <- (gapminder %>%
  filter(country == 'Greece' | country == 'Austria' | country == 'Philippines', year == 1970:1979))
gapminder1.1
```

## 1.2

Show variables country and gdpPercap only from gapminder1.1:
```{r}
gapminder1.1 %>%
  select(country, gdpPercap)
```

## 1.3

To filter gapminder to all rows that have a drop in life expectancy a new temporary variable of countrylag containing country values shifted one row down is made.
A new variable of increase in life expectancy is made from the difference in life expectancy between each year in order of country and year. Country and countrylag values are compared.  If they are not equal then its value is replaced with 0.  Increase in life expectancy is filtered for drop in life expectancy by filtering for negative values only. Countrylag is removed from the table.

```{r}
gapminder %>%
  mutate(countrylag = lag(country, order_by = country & year)) %>%
  mutate(increaseLifeExp = lag(diff(gapminder$lifeExp), order_by = country & year)) %>%
  mutate(increaseLifeExp = if_else( country == countrylag, increaseLifeExp, 0)) %>%
  filter(increaseLifeExp < 0) %>%
  select(-countrylag)
```


## 1.4

To filter gapminder to find the rows with the 3 largest and 3 smallest GDP per capita you can:

Find the number of rows in gapminder:
```{r}
nrow(gapminder)
```
 
Arrange gapminder in descending order for GDP per capita, then select top 3 rows and bottom 3 rows:

```{r}
(gapminder %>%
  arrange(desc(gdpPercap)))[c(1:3, 1702:1704), ]
```

## 1.5

Scatterplot of Canada's life expectancy vs. GDP per capita:

```{r}
ggplot(gapminder %>%
  filter( country == 'Canada'), aes(gdpPercap, lifeExp)) +
  geom_point(colour = 'blue') +
  scale_x_log10('log(gdpPercap)') 
```

## 2

An exploration of continent and population variables.  Here is the data:


```{r}
gapminder %>%
          select(continent, pop)
```


Here is a summary of the data for continent and population variables:
```{r}
summary(gapminder %>%
          select(continent, pop))
```
```{r}
1.319e+09 - 6.001e+04
```

The possible values for continent are Africa, Americas, Asia, Europe, and Oceania.  Population has a maximum value of 1.319e+09 and a minimum value of 6.001e+04 with a range of 1318939990. 

The following graph represents the population of each continent from 1951-2007:

```{r}
ggplot(gapminder, aes(continent, pop)) +
  geom_col() 
```
```{r}
range(gapminder$year)
```
Asia has the largest population, while Oceania has the smallest population of the 5 categories.  Africa, Americas, and Europe have somewhat similar population levels relative to Asia and Oceania.


## 3


Exploration of Loblolly pine trees.  Here is a scatterplot of height (ft) vs. age (yr):

```{r}
ggplot(Loblolly, aes(age, height)) +
         geom_point(colour = 'magenta')
```

Here is a bar graph of height (ft) vs. seed source:

```{r}
ggplot(Loblolly, aes(Seed, height)) +
         geom_col()
```
Seed 305 appears to have the largest height.


