---
title: "hw01_presentation"
output: ioslides_presentation
---


## Loading the Dataset

R has a dataset called gapminder, which the following code loads:

```{r load}
library(gapminder)
```


## Variables

Here are the variables for gapminder after running the following code:

```{r, results='hide'}
names(gapminder)
```
country, continent, year, lifeExp, pop, gdpPercap

## Rows

Here are the number of rows in gapminder after running the following code:

```{r}
nrow(gapminder)
```

## Dataset Summary

Here is a summary of statistics of gapminder:

```{r, results='hide'}
summary(gapminder)
```
country | continent  |  year  
---------- | ------------ | ------------
 Afghanistan:  12  | Africa  :624 |  Min.   :1952  
 Albania    :  12 |  Americas:300 |  1st Qu.:1966   
 Algeria    :  12 |  Asia    :396 |  Median :1980    
 Angola     :  12 |  Europe  :360 |  Mean   :1980  
 Argentina  :  12 |  Oceania : 24 |  3rd Qu.:1993   
 Australia  :  12 |               |  Max.   :2007  
 

## Dataset Summary Continued

lifeExp  |         pop       |      gdpPercap  
--------- | --------------- | ----------------
 Min.   :23.60  | Min.   :6.001e+04  | Min.   :   241.2
 1st Qu.:48.20  | 1st Qu.:2.794e+06 |  1st Qu.:  1202.1 
 Median :60.71  | Median :7.024e+06 |  Median :  3531.8  
 Mean   :59.47  | Mean   :2.960e+07 |  Mean   :  7215.3 
 3rd Qu.:70.85  | 3rd Qu.:1.959e+07 |  3rd Qu.:  9325.5 
 Max.   :82.60  | Max.   :1.319e+09 |  Max.   :113523.1  
                                                            

## Countries

Here is a list of countries within the gapminder dataset using the following code:

```{r, results='hide'}
(unique(gapminder$country))
```
Afghanistan              Albania                  Algeria             Angola                   Argentina                Australia               Austria                  Bahrain                  Bangladesh            Belgium                  Benin                    Bolivia               Bosnia and Herzegovina   Botswana                 Brazil                Bulgaria                 Burkina Faso             Burundi                 Cambodia                 Cameroon                 Canada                Central African Republic Chad                     Chile                   China                    Colombia                 Comoros                 Congo, Dem. Rep.         Congo, Rep.              Costa Rica              Cote d'Ivoire            Croatia                  Cuba                  Czech Republic           Denmark                  Djibouti              Dominican Republic   Ecuador                  Egypt                   El Salvador              Equatorial Guinea        Eritrea                 Ethiopia                 Finland                  France                  Gabon                    Gambia                   Germany               Ghana                    Greece                   Guatemala               Guinea                   Guinea-Bissau            Haiti                    Honduras                 Hong Kong, China         Hungary                 Iceland                  India                    Indonesia               Iran                     Iraq                     Ireland                  Israel                   Italy                    Jamaica                  Japan                    Jordan                   Kenya                   Korea, Dem. Rep.         Korea, Rep.              
 
## Countries Continued

Kuwait                   Lebanon                  Lesotho                  Liberia                  Libya                    Madagascar               Malawi                   Malaysia                 Mali                     Mauritania               Mauritius                Mexico                   Mongolia                 Montenegro               Morocco                  Mozambique              Myanmar                  Namibia                  Nepal                    Netherlands              New Zealand              Nicaragua                Niger                    Nigeria                  Norway                   Oman                     Pakistan                 Panama                   Paraguay                 Peru                     Philippines              Poland                   Portugal                 Puerto Rico              Reunion                  Romania                  Rwanda                   Sao Tome and Principe    Saudi Arabia             Senegal                  Serbia                   Sierra Leone             Singapore                Slovak Republic          Slovenia                 Somalia                 South Africa             Spain                    Sri Lanka               Sudan                    Swaziland                Sweden                   Switzerland              Syria                    Taiwan                 Tanzania                 Thailand                 Togo                     Trinidad and Tobago      Tunisia                  Turkey                   Uganda                   United Kingdom           United States            Uruguay                  Venezuela                Vietnam                  West Bank and Gaza       Yemen, Rep.              Zambia                   Zimbabwe                